On Etiology of Cybersecurity

### 2020-07-13 reports/paper_0/main.tex
Luca

@@

\begin{abstract}
The objective of this research is the development of a theory that defines (all
	and only) the possible insecurity and security configurations of any
	abstract system. The theory is structured upon other theories that
	defines how a component of a system can be abstracted into an agent,
	defining how agents can be formalized (both syntactically and
	semantically) to describe an abstract system, such as a graph. The core
	theories are the epistemological definition of knowledge, Beliefs, and
	Information, the Assertion-Belief-Fact framework, and mereotopology.
	We implemented a formal theory (of axioms) of a mereotopology, and of
	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
	uses the Z3 SMT solver. The results show that a single component (i.e.
	agent) of an abstract system has a definite number of  different
	insecurity configurations (e.g. 53 using RCC5 over a topological
	structure) and only 1 secure (i.e.  expected) configuration. The
	configurations are reported as models satisfying the abstract system
	semantics. We use these results to predict all possible security
	weaknesses of a system, instead of relying on databases of known
	weaknesses, and to lay the foundations of a scientific theory of
	cybersecurity.  We show a concrete applications of our theory to the
	risk assessment of an ad-hoc system.
\end{abstract}

@@

\section{Introduction}\label{sec:intro}
A scientific theory is an explanation of a phenomenon such that the explanation
follows the scientific method. The scientific method is an \emph{empirical}
method that aims at mitigating potential fallacies in theories.  K. Popper
famously argued (e.g. in \autocite{Popper1959logic}) that a scientific theory
can never been verified but only falsified, that a theory should not be
conceived by using the principle of induction\footnote{A.  Einstein wrote to K.
Popper: ``[...] and I think (like you, by the way) that theory cannot be
fabricated out of the results of observation, but that it can only be
invented.''\autocite{Popper1959logic}}, and that empirical experiments should
be considered as the only evidence to support the non-falseness of a scientific
theory. If we consider a natural (real-world) phenomenon, such as gravity, we
can think of a scientific theory as an \emph{abstract} mathematical formula (or
a set of formulas) that axiomatically define a certain (natural) phenomenon in
a specific algebraic structure. While such a theory can be proved to be correct
within the mathematical setting it has been conceived in (e.g. with respect to
other axioms of the algebraic structure where it has been postulated, or with
respect to other background theories), no proof (for soundness or completeness)
can be carried out to verify that the theory correctly abstracts  the
phenomenon it describes. Similarly, the results of the application of a
scientific theory (i.e. its predictions) can be verified in the abstract
mathematical structure but how can we be sure that the theory correctly
characterize the phenomenon it describes? With his famous ``demarcation
principle'', K. Popper draw a line between scientific and pseudo-scientific
theories by proposing that \emph{any} scientific theory can be considered
correct as long as no empirical experiment falsify it. For example, the theory
of gravity, in its mathematical setting, can predict how object attracts with
each other, and those predictions can be tested by empirical experiment. If the
theory correctly abstracts the phenomenon no empirical experiment should
falsify the predictions of that theory.

In \autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
``an asymmetry in computer security'', which he defines as follows: ``Things
can be declared insecure by observation, but not the reverse. There is no
observation that allows us to declare an arbitrary system or technique
secure''. His intuition is that there is no scientific theory that can predict
the security of a system, nor a theory that can predict all possible
insecurity of a system (which, by negation it would become a theory on security). 
Herley then uses this argument to show that ``claims that any measure
is necessary for security are empirically unfalsifiable''. Given that, any
theory which is not falsifiable by an empirical experiment is well
known\footnote{``A theory which is not refutable by any conceivable event is
nonscientific. Irrefutability is not a virtue of a theory (as people often
think) but a vice.'' -- Karl Popper, Conjectures and
Refutations \autocite{popper1962conjectures}} to be nonscientific (i.e.
unfalsifiability is a fallacy of a theory), Herley concludes that there is no
scientific theory on cybersecurity; which means that cybersecurity lays in the
realm of pseudo-sciences \autocite{Herley2016usenixvideo}.  Herley, e.g.
in \autocite{Herley2017justifying}, discusses the implications of a
nonscientific approach to cybersecurity, and highlights the tremendous impact on
all the scientific research and engineering of systems; leading often to
terrorism and wars, and wasting of resources in useless protections or
overspending.  While the criticism is investigated
in \autocite{Herley2016unfalsifiability}, no solution is provided nor envisioned.  On the
contrary, the goal of this work is to lay the foundations of a
scientific cybersecurity theory.

We consider the problem raised by Herley not confined to ``computer security''
but to any abstract system (so that our theory may hold for any sound
implementation such as networks, mechanical, cyber, or cyber-physical system,
or even a single computer or a single device such as an hard-drive).  There is

@@

seek to clarify before following (as we agree) the scientific path draw by

@@

security properties verified).  For example, in the protocol verification

@@

(e.g. ) used for the engineering of systems starts with the requirements

@@

that may generate security vulnerabilities later in the development process. An

@@

the authors wrote\autocite{krak2020web}: ``our attacks do not violate the security properties proven

@@

proofs do not model key installation.''.

@@

non-evidently apply to a general security theory.  As an example, the so called

@@

protocol. This, in turn, creates a false sense of security since requires

@@

is verified. More specifically, for the formal security verification of a

@@

often done when verifying the security of authentication protocols), (ii) the

@@

of sessions such as \autocite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
the peers (e.g.  in the ASLan++ language \autocite{Oheimb2010aslan++}), and (iv)

@@

\autocite{Blanchet2017symbolic}).  While many tools and theories improves those 

@@

approach (if any). More importantly, some of the choices made by the

@@

is suppose to be perfect, without any exploitable flaw, and so the only way for

@@

\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
and assuming that no violation to any security property is done after message

@@

determines that the flaws related to the dishonest impersonation of the Server
may or may not be considered in the verification process.  This choice has

@@

non-falsifiable theory of system security

>>>>

\newcommand{\fixn}[2]{{\bf FIX}\footnote{#1: #2}}

@@

\begin{abstract}
\fixn{Luca}{Per il titolo, meglio: ``On the Etiology of Cybersecurity'' oppure ``The Etiology of Cybersecurity''.
}
The objective of this research is the development of a theory that defines (all
	and only) the possible insecurity and security configurations of any
	abstract system. The theory is structured upon other theories that
	define how a component of a system can be abstracted into an agent,
	defining how agents can be formalized (both syntactically and
	semantically) to describe an abstract system, such as a graph.\fixn{Luca}{la frase subordinata ``defining...'' non si capisce a cosa si riferisca. Troppi ``define'' e troppo lunga la frase.} The core
	theories are the epistemological definition of knowledge, Beliefs, and
	Information, the Assertion-Belief-Fact framework, and mereotopology.
	We\fixn{Luca}{Non si capisce come questa frase si colleghi a quanto sopra. E se e come questa formal theory sia related a sopra} implemented a formal theory (of axioms) of a mereotopology, and of
	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
	uses the Z3 SMT solver. The results\fixn{Luca}{The results of what?} show that a single component (i.e.
	agent) of an abstract system has a definite number of  different
	insecurity configurations (e.g. 53 using RCC5 over a topological
	structure) and only 1 secure (i.e.  expected) configuration. The
	configurations are reported as models satisfying the abstract system
	semantics.\fixn{Luca}{Non si capisce la relazione tra agent e abstract system semantics} We use these results to predict all possible security
	weaknesses of a system, instead of relying on databases of known
	weaknesses, and to lay the foundations of a scientific theory of
	cybersecurity.  We show a concrete applications of our theory for the risk assessment of an ad-hoc system.\fixn{Luca}{Io continuo a pensare che non si capisca bene cosa stiate facendo. \`E ancora scritto in maniera troppo complicata e vengono dette troppe cose. Io accorcerei e semplificherei.}
\end{abstract}

@@

\section{Introduction}\label{sec:intro}
\paragraph{Context and motivation} 
A \emph{scientific theory} is an explanation of a phenomenon such that the explanation
follows the scientific method. The \emph{scientific method} is an \emph{empirical}
method that aims at mitigating potential fallacies in theories.  Karl Popper
famously argued (e.g. in \autocite{Popper1959logic}) that a scientific theory
can never be verified but only falsified, that a theory should not be
conceived by using the principle of induction\footnote{Albert  Einstein wrote to 
Popper: ``[...] and I think (like you, by the way) that theory cannot be
fabricated out of the results of observation, but that it can only be
invented.''~\autocite{Popper1959logic}}, and that empirical experiments should
be considered as the only evidence to support the non-falseness of a scientific
theory.

If we consider a natural (real-world) phenomenon, such as gravity, we
can think of a scientific theory as an \emph{abstract} mathematical formula (or
a set of formulas) that defines axiomatically a certain (natural) phenomenon in
a specific algebraic structure. While such a theory can be proved to be correct
within the mathematical setting it has been conceived in (e.g. with respect to
other axioms of the algebraic structure where it has been postulated, or with
respect to other background theories), no proof (for soundness or completeness)
can be carried out to verify that the theory correctly abstracts the
phenomenon it describes.\fixn{Luca}{Troppi correct. Meglio chiarire bene cosa sono correctness, soundness e completeness} Similarly, the results of the application of a
scientific theory (i.e. its predictions) can be verified in the abstract
mathematical structure but how can we be sure that the theory correctly
characterizes the phenomenon it describes?

Popper's famous ``demarcation
principle'' draws a line between scientific and pseudo-scientific
theories by proposing that \emph{any} scientific theory can be considered
correct as long as no empirical experiment falsify it. For example, the theory
of gravity, in its mathematical setting, can predict how objects attract
each other, and those predictions can be tested by empirical experiment. If the
theory correctly abstracts the phenomenon, no empirical experiment should
falsify the predictions of that theory.

In \autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
``an asymmetry in computer security'', which he defines as follows: ``Things
can be declared insecure by observation, but not the reverse. There is no
observation that allows us to declare an arbitrary system or technique
secure''. His intuition is that there is no scientific theory that can predict
the security of a system, nor a theory that can predict all possible
insecurities of a system (which, by negation would become a theory on security). 
Herley then uses this argument to show that ``claims that any measure
is necessary for security are empirically unfalsifiable''.\fixn{Luca}{Whay if I have a system that I have proved to be secure under a set of $n$ assumptions. Now I remove those $n$ assumptions one by one and run the proof again. What if the proof works whenever I remove one of the first $n-1$ assumptions, but fails (only) if I remove the $n$th assumption. Would the counterexample, which can be applied empirically, not be evidence that the assumption removed was a necessary measure?} Given that any
theory that is not falsifiable by an empirical experiment is well
known\footnote{``A theory which is not refutable by any conceivable event is
nonscientific. Irrefutability is not a virtue of a theory (as people often
think) but a vice.'' -- Karl Popper, Conjectures and
Refutations~\autocite{popper1962conjectures}.} to be nonscientific (i.e.
unfalsifiability is a fallacy of a theory), Herley concludes that there is no
scientific theory on cybersecurity; which means that cybersecurity lays in the
realm of pseudo-sciences~\autocite{Herley2016usenixvideo}.\fixn{Luca}{Theory ON cybersecurity or theory OF cybersecurity. A me viene da dire OF}  Herley, e.g.
in \autocite{Herley2017justifying}, discusses the implications of a
nonscientific approach to cybersecurity, and highlights the tremendous impact on
all the scientific research and engineering of systems; leading\fixn{Luca}{Questo suona un po' esagerato, Meglio chiarire.} often to
terrorism and wars, and wasting of resources in useless protections or
overspending.  While the criticism is investigated
in \autocite{Herley2016unfalsifiability}, no solution is provided nor envisioned.

\paragraph{Contributions} 
The goal of this paper is to lay the foundations of a
scientific cybersecurity theory. 
We consider the problem raised by Herley not confined to ``computer security''
but to any abstract system (so that our theory may hold for any sound
implementation such as networks, mechanical, cyber, or cyber-physical system,
or even a single computer or a single device such as a hard-drive).\fixn{Luca}{Non si capisce ``any'' system. Quindi anche un sistema che non ha nulla a che fare con security? La frase ``not confined to... but to... '' non \`e grammaticamente corretta. Il verbo non pu\`o rimanere ``confined'', casomai ``not confined to ``computer security''
but rather we consider any abstract system'' o una cosa del genere}

There is

@@

seek to clarify before following (as we agree) the scientific path drawn by

@@

security properties verified).\fixn{Luca}{Questa frase non la capisco. Innanzitutto non capisco la differenza tra cybersecurity e security. E poi cybersecurity non \`e una property ma un set of properties. E non si capisce come si passa da un system ad un protocol... sembrano due frasi diverse. Io le spezzerei, parlando del problema in generale, e poi dicendo, As a concrete example, consider the case of security protocols...}  For example, in the protocol verification

@@

(e.g. \fixn{Luca}{manca ref}) used for the engineering of systems start with the requirements

@@

that may generate security vulnerabilities later in the development process.\fixn{Luca}{Qui si cercano di dire troppe cose. Improvvisamente, mentre si parlare di security definition, si parla del software life-cycle. Va spiegato molto meglio (e in maniera pi\`u semplice.) Inoltre, non \`e proprio corretto. Che dire di model-based testing in cui si fa analysis della specifica e dell'implementazione?} An

@@

the authors\fixn{Luca}{The authors of what? Solo Vanhoef?} wrote~\autocite{krak2020web}: ``our attacks do not violate the security properties proven

@@

proofs do not model key installation.''\fixn{Luca}{Questo mi sembra un modo confuso e un po' specioso di dire un'ovviet\`a: i risultati di un'analisi formale sono buoni tanto quanto il modello. Se il modello non modella una cosa, allora non si possono dedurre conclusioni su quella cosa. Questo \`e un problema generico della modellizzazione di qualunque cosa, non solo di security. E non direi che questo sia un esempio infamous, visto che ne esistono molti altri molto pi\`u importanti.}

@@

non-evidently apply to a general security theory.\fixn{Luca}{Non capisco ``which
non-evidently apply to a general security theory''}  As an example, the so called

@@

protocol.\fix{Luca}{Io lo direi diversamente: ``Dolev-Yao attacker model'' is an abstract model of an attacker that can be used for the analysis of protocol specifications} This, in turn, creates a false sense of security since it requires

@@

is verified.\fixn{Luca}{Frase scorretta grammaticamente ma anche molto confusa. Io direi: a proof of security of the protocol specification under the Dolev-Yao attacker model may instill a false sense of security of the whole protocol, whereas we have only verified it for the speicification. O una cosa del genere. In generale, ci sono troppo salti logici nei claims, dettati dalla voglia di dire troppe cose subito.} More specifically, for the formal security verification of a

@@

often done when verifying the security of authentication protocols),fixn{Luca}{Nowadays, most tools don't remove these} (ii) the

@@

of sessions such as \autocite{Escobar2007maudenpa}),\fixn{Luca}{Nowadays, most tools work for infinite sessions: Maude-NPA, Tamarin, ProVerif, OFMC.} (iii) honesty/dishonesty of
the peers (e.g.  in the ASLan++ language \autocite{Oheimb2010aslan++}),\fixn{Luca}{Nowadays, most tools allow for easy switch of the resulting sessions} and (iv)

@@

\autocite{Blanchet2017symbolic}).\fixn{Luca}{This is the only claim that, more or less, still holds, even though more and more tools do cryptoanalysis. The good news is that, I believe, you do not need (i)-(iv) at all, but you can simply say that by the nature of modeling a specification, you are abstracting away certain things (e.g., cryptography, actual message content, PKI, etc.), and the vulnerabilities might be hidden there} While many tools and theories improved those 

@@

approach (if any).\fixn{Luca}{I don't think that this is important. In many fields of research you don't have a single definitive approach and it is good that it is so. I would remove this sentence, but only focus on the next point, which is the only important one of this discussion of protocols} More importantly, some of the choices made by the

@@

is supposed to be perfect, without any exploitable flaw, and so the only way for

@@

\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}\fixn{Luca}{I would not quote your own paper for something for which there are hundreds of other equivalent quotes. Just say ``under the perfect
cryptography assumption of the Dolev-Yao attacker model''}
and assuming that no violation to any security property occurs after message

@@

determines that the flaws related to the dishonest impersonification of the Server
may or may not be considered in the verification process.  This choice has a

@@

non-falsifiable theory of system security.\fixn{Luca}{Frase troppo complicata, alla fine di un paragrafo alquanto confuso e ridondante. Secondo me tutto questo discorso si pu\`o ridurre a tre frasi, come ho suggerito sopra.'}
